---
title: "lucas_county_local"
format: html
editor: visual
---

```{r}
library(Jmisc)
library(sf)
library(tidyverse)
library(ggplot2)
library(sp)
library(spdep)
library(spatialreg)
library(spData)
library(modelsummary)
library(flextable)
library(knitr)
# Options
options(scipen = 7)
```

# Local Phenomena

```{r}
# objects defined in intro.qmd
load("lucas.RData")
```

## Generate neighbour structure

Model used

```{r}
cat(hmod1)
```

```{r}
# lucas_w was generated in the into.qmd file
lucas_w
```

## Moran Plot

```{r}
moran.plot(log(lucas$price), listw = lucas_w, labels = FALSE, pch = 20, cex = 0.3)
```

## Moran Plot; influential obs.

-   If we want to see which obs.
    are the most influential we can change labels to TRUE

-   The row number for each influential obs.
    will then be plotted in the figure.

```{r}
moran.plot(lucas$age_sold, listw = lucas_w, 
           labels = TRUE, pch = 20, cex = 0.3)
```

## Approximate profile-likelihood estimator (APLE)

Approximate profile-likelihood estimator (APLE), see Li, Calder, and Cressie ([2007](https://msb205.netlify.app/spatmetrics/lucascounty/local#/references)) and Li, Calder, and Cressie ([2012](https://msb205.netlify.app/spatmetrics/lucascounty/local#/references)).

```{r}
# Calculate APLE for the variables
lucas %>% 
  select(price, age_sold, TLA, lotsize, rooms, beds) %>% 
  st_set_geometry(., NULL) %>% 
  demean() %>% 
  as_tibble() %>% 
  map(aple,listw = lucas_w) |> 
  unlist()
```

## Conclusion APLE

-   We see that the strength of positive spatial dependence is greatest for `price`, `TLA` and `lotsize` and lowest for number of bedrooms (`beds`) in the dwellings.

-   APLE, which may be viewed somewhat like a correlation coefficient between the values observed for each state and the average values for neighbouring states.([R. Bivand, Millo, and Piras 2021a](https://msb205.netlify.app/spatmetrics/lucascounty/local#/references))

## Local phenomena; model list

-   SDEM,

    -   `SDEM <- errorsarlm(hmod1, data = lucas, listw = lucas_w,  Durbin = TRUE, na.action = na.omit)`

-   SEM,

    -   `SEM <- errorsarlm(hmod1, data = lucas, listw = lucas_w,  Durbin = FALSE, na.action = na.omit)`

-   SLX,

    -   `SLX <- lmSLX(hmod1, data = lucas, listw = lucas_w, Durbin = TRUE, na.action = na.omit)`

-   lm (OLS)

    -   `lm1 <- lm()`

## Local, start with SDEM

-   Some of these models are quite expensive to compute, so we cache the results

    -   It should automagically be recomputed if something change

-   But caching is complex so we should check that the models really are recomputed if we change the data or model

-   Note the use of the option `Durbin` to choose the kind of model.

    -   Used to be done by the options `type` and `etype` until recently.

```{r}
SDEM <- errorsarlm(hmod1, data = lucas, listw = lucas_w, 
                   Durbin = TRUE, na.action = na.omit)
```

```{r}
summary(impacts(SDEM), zstats = TRUE)
```

## Local, SEM

```{r}
# No spatial effects in the explanatory variables
SEM <- errorsarlm(hmod1, data = lucas, listw = lucas_w, 
                  Durbin = FALSE, na.action = na.omit)
```

```{r}
# normal summary, no impacts
summary(SEM)
```
